---
output:
  html_document:
    code_folding: hide
  html_notebook: default
  pdf_document: default
---
# UK to EU exports - monthly totals as at January 2021

This uses UK Office for National Statistics data published on  12th March 2021, available at the link below:

https://www.ons.gov.uk/economy/nationalaccounts/balanceofpayments/datasets/uktradeallcountriesseasonallyadjusted


```{r, warning = F, message = F}
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(tidyquant)

```


## Load and clean data

This uses data on UK exports from this Excel file:
https://www.ons.gov.uk/file?uri=/economy/nationalaccounts/balanceofpayments/datasets/uktradeallcountriesseasonallyadjusted/current/allcountriesjan2021final.xlsx

I have copied the EU data from the *Monthly Exports* tab into a separate csv saved in the `data_raw` folder, which is the one I'm loading in here.

The full Excel file is also saved for reference in the `data_raw` folder.

```{r, warning = F, message = F}
# Read in exports data
uk_to_eu_exports <- read_csv(here("data_raw/exports.csv"))

# Pivot longer - convert month columns to rows
uk_to_eu_exports <- uk_to_eu_exports %>% 
  pivot_longer(cols = "01/01/1997":"01/01/2021",
               names_to = "month",
               values_to = "exports_pounds_million")

# Convert month column to date format
uk_to_eu_exports <- uk_to_eu_exports %>% 
  mutate(month = dmy(month))

# Add year_month column
uk_to_eu_exports <- uk_to_eu_exports %>%
mutate(month_year = zoo::as.yearmon(month))

# Reorder data based on exports in Dec 2020 (highest to lowest by value)
uk_to_eu_exports <- uk_to_eu_exports %>%
mutate(country_name = factor
(country_name, levels =
c("Total EU",
"Germany",
"Ireland",
"Netherlands",
"France",
"Belgium",
"Italy",
"Spain",
"Poland",
"Sweden",
"Denmark",
"Czech Republic",
"Portugal",
"Austria",
"Hungary",
"Romania",
"Finland",
"Greece",
"Slovakia",
"Cyprus",
"Malta",
"Lithuania",
"Latvia",
"Bulgaria",
"Estonia",
"Slovenia",
"Luxembourg",
"Croatia"))) 
```



## Visualise the data - Dec 2019 to Jan 2021

### Total exports to EU
```{r}
uk_to_eu_exports %>% 
filter(month_year >= "2019-12") %>% 
filter(country_name == "Total EU") %>% 
ggplot()+
  aes(x = month, y = exports_pounds_million) +
  geom_line(size  = 2, colour = "steel blue") +
  theme_minimal() + 
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x = element_text(angle=45,hjust=1)) +
  theme(plot.subtitle = element_text(size = 8)) +
  theme(axis.text = element_text(size = 7)) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(axis.title = element_text(size = 8)) +
   theme(legend.position="bottom") +
  labs(
    title = "UK Trade in Goods - Total for All EU Countries - Monthly Exports",
    subtitle = "Source: Office for National Statistics \n",
    x = "",
    y = "£ million, seasonally adjusted, current price \n",
    colour = ""
  ) 


```


### Exports to selected EU countries
```{r}
uk_to_eu_exports %>% 
filter(month_year >= "2019-12") %>% 
filter(country_name %in% c(
"Germany",
"Ireland",
"Netherlands",
"France",
"Belgium")) %>% 
ggplot()+
  aes(x = month, y = exports_pounds_million, group  = country_name, colour = country_name) +
  geom_line(size  = 2) +
  theme_minimal() + 
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x = element_text(angle=45,hjust=1)) +
  theme(plot.subtitle = element_text(size = 8)) +
  theme(axis.text = element_text(size = 7)) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(axis.title = element_text(size = 8)) +
   theme(legend.position="bottom") +
  labs(
    title = "UK Trade in Goods - Selected EU Countries - Monthly Exports",
    subtitle = "Source: Office for National Statistics \n",
    x = "",
    y = "£ million, seasonally adjusted, current price \n",
    colour = ""
  ) 


```

